---
title: "p8105_hw1_slv2129"
author: "Krishna Vemulapalli"
date: "20/09/2020"
output: github_document
---

# Problem 1

Installing tidyverse library to call the tibble function

```{r, echo = false, message = false}
library(tidyverse)
```

Creating a data frame consisting of the following:
  * a random sample of size 10 from a standard Normal distribution
  * a logical vector indicating whether elements of the sample are greater than 0
  * a character vector of length 10
  * a factor vector of length 10, with 3 different factor “levels”

```{r, echo = false}
nd_df = tibble(
  rand_samp = rnorm(10),
  log_vec = rand_samp > 0,
  char_vec = c("the", "and", "in", "tree", "wood", "bird", "fish", "food", "rock", "stone"),
  fac_vec = factor(c("good", "bad", "average", "good", "bad", "average", "good", "bad", "average",   "good"))
)

 nd_df
```

Taking mean of each variable in nd_df

```{r, ech0 = false}
mean(nd_df$rand_samp)

mean(nd_df$log_vec)

mean(nd_df$char_vec)

mean(nd_df$fac_vec)
```

From the above we can deduce one can find the mean for numeric and logical vectors. Whereas, the mean can't be calculated for character and factor data type vectors. 

Converting the logical, character, and factor data type variables to numeric

```{r, echo = false, results = hide, warning = false}
log_num = as.numeric(nd_df$log_vec)

char_num = as.numeric(nd_df$char_vec)

fac_num = as.numeric(nd_df$fac_vec)
```

Both factor and logical variables got number coded. The "good", "bad" and "average" have been replaced with 3,2 and 1 respectively for the factor variable. In the case of logical vector, "true" (condition satisfied) has been coded as 1 and "false" has been coded as 0. All the components of character variable have been replaced with "NA". This explains why the mean function can only be applied to numeric and logical variable. However, the factor variable can't perform numeric operations despite the numerical representation. 

The following actions will be performed in the chuck of code below:
  * convert the logical vector to numeric, and multiply the random sample by the result
  * convert the logical vector to a factor, and multiply the random sample by the result
  * convert the logical vector to a factor and then convert the result to numeric, and multiply the        random sample by the result
  
```{r, echo = false, warning = false}
num_log = as.numeric(nd_df$log_vec) * nd_df$rand_samp

fac_log = as.factor(nd_df$log_vec) * nd_df$rand_samp

fac_log_num = as.numeric(as.factor(nd_df$log_vec)) * nd_df$rand_samp

#Printing the values of the resultant vectors:#

num_log

fac_log

fac_log_num

```

